---
title: "EDS 221 Day 7 Morning interactive"
format: html
---

```{r}
library(tidyverse)
library(palmerpenguins)
library(lubridate)
```


# Data Wrangling refresher 

### Refresher 1
- Only include penguins at Biscoe and Dream islands
- Remove the year and sex variables
- Add a new column called body_mass_kg, with penguin mass converted from grams to kilograms
- Rename the island variable to location
```{r}
# load in penguins dataset 
penguins_1 <- penguins %>%
  filter(island == "Biscoe" | island == "Dream") %>%
  select(-sex, -year) %>% 
  mutate(body_mass_kg = body_mass_g * 0.001) %>% 
  rename(location = island)
```


### Refresher 2
- Limit to only Adelie penguins
- Remove any observations where flipper_length_mm is NA (hint: !is.na())
- Group the data by sex
- Find the mean (mean()), standard deviation (sd) and sample size (n) of flipper_length_mm for male and female Adelie penguins, returned in a nice summary table
```{r}
penguins_2 <- penguins %>% 
  filter(species == "Adelie") %>% 
  filter(!is.na(flipper_length_mm) & !is.na(sex)) %>% 
  group_by(sex) %>% 
  summarize(count = n(), mean_lengths = mean(flipper_length_mm), sd_lenght = sd(flipper_length_mm))

```




